---
title: "Cars are Cool"
author: "Sam Kerns"
date: "4/26/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(kableExtra)
library(assertr)
```

### Using this dashboard

Cars are cool - or at least I think so. To get a sense of what you're getting into when you purchase your next used vehicle, I have created a dashboard that allows you to look at the various horsepower ratings of car models by their manufacturer.

```{r, include = F, results = "hide"}
cars <- mtcars
head(cars)

```

Now that we've loaded up the dataset and taken a quick peek at the data, let's get a sense of the average horsepower per number of cylinders in the car

```{r, results = "html"}

avg_hp <- cars %>%
        group_by(cyl) %>%
        summarize(avg_hp = round(mean(hp, na.rm = T), 2))

knitr::kable(avg_hp) %>%
  kableExtra::kable_styling()
```

Cool table...but what about a graph?

```{r, results = "html"}
ggplot(data = avg_hp, aes(x = cyl, y = avg_hp)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = avg_hp), vjust = 1.5, colour = "white") +
  scale_x_continuous(breaks = c(4, 6, 8)) +
  theme_minimal()
```

Now that we've done a little look into the data, let's create a tabbed output of the power output per cylinder based on region of origin. First, we'll need to extract the make from the rownames and then we'll convert those to American, European, and Japanese cars:

```{r}
##Making lists for regional car manufacturers
american_cars <- c("Hornet", "Valiant", "Duster", "Cadillac", "Lincoln", "Chrysler", "AMC", "Ford", "Pontiac", "Camaro", "Dodge")
euro_cars <- c("Merc", "Fiat", "Porsche", "Lotus", "Ferrari", "Maserati", "Volvo")
japanese_cars <- c("Mazda", "Datsun", "Toyota", "Honda")

##Creating a region variable based on above lists
cars <- cars %>%
  mutate(make = str_extract(row.names(cars), "^[A-Za-z]+")) %>%
  mutate(region = case_when(make %in% american_cars ~ "American",
                            make %in% euro_cars ~ "European",
                            make %in% japanese_cars ~ "Japanese")) %>%
  verify(!is.na(region))
```

Now, let's go ahead and tab out the output so we can see the power by number of cylinders for each of the regions from which the cars in this dataset are manufactured!

## Average Horsepower by Number of Cylinders and Region {.tabset}

```{r, include = T, results = "asis"}
##Reshaping the data to prepare for ggplot
cars_plot <- cars %>%
  group_by(region, cyl) %>%
  summarise(Avg_HP = round(mean(hp, na.rm = T), 2))

##Iterate over regions
for (r in c("American", "European", "Japanese")) {
  
  ##Tab header
  cat(paste0("\n\n", "### ", r, "\n\n"))
  
  ##Reducing dataset
  cars_reduce <- cars_plot %>%
    filter(region == r)
  
  ##Horsepower by cylinder count
  fig <-
    ggplot(cars_reduce, aes(x = cyl, y = Avg_HP)) +
    geom_col(color = "#046B5C", fill = "#046B5C") +
    geom_text(aes(label = Avg_HP), vjust = 1.5, colour = "white") +
    xlab("Number of Cylinders") +
    ylab("Mean Horsepower") +
    theme_minimal()
  
  ##Printing the figure
  print(fig)
  
  ##Adding the figure to the tabset
  cat("\n\n")
}
```
